A Wait-free Queue with Polylogarithmic Step Complexity

### 2023-01-06 podc_paper/algorithm-2.tex
finished draft of code for GC

\Function{void}{Propagate}{\typ{Node} \var{v}}
    \State \linecomment{propagate blocks from \var{v}'s children to root}
    \If{\bf{not} \Call{Refresh}{\var{v}}} \label{firstRefreshGC}  \hfill \com\ double refresh
        \State \Call{Refresh}{\var{v}} \label{secondRefreshGC}
    \EndIf
    \If{\var{v} \bf{is not} \var{root}} \hfill \com\ recurse up tree
        \State \Call{Propagate}{\var{v}.\fld{parent}}
    \EndIf
\EndFunction{Propagate}

@@

\gc{    
	\If{ \var{v} is root and $\var{h} \mbox{ mod } p^2=0$}
	\State \var{max} \assign 0
	\For {$k \assign 1..p$}
		\State $\var{max} \assign \max(\var{max}, \var{last}[k])$
	\EndFor 
	\State \Call{Help()}{}
	\State \Call{Discard}{$\var{root}, \var{max}-1$}
	\EndIf
	\State
}

@@

\end{algorithmic}
\end{minipage}
\caption{Queue implementation.\label{pseudocode1GC}}
\end{figure}

@@

\gc{
\Function{void}{Help}{}
\State \bf{for} \var{i} \assign\ 1 to $p$
\State h=leafOfProcess[\var{i}]
\If{leafOfProcess[\var{i}].blocks[h].num\sub{deq}==1 and leafOfProcess[\var{i}].IndexDequeue(h,1)!=null}
\State <b, i> \assign\ \Call{IndexDequeue}{leafOfProcess[\var{i}], h, 1}
\State output \assign\ \Call{FindResponse}{b, i} 
\State leafOfProcess[\var{i}].blocks[h].response \assign\ output
\EndIf
\State \bf{end for}
\EndFunction{Help}

\Function{void}{Discard}{\typ{Node} \var{v}, \typ{int} \var{i}}
    \State $T \assign \var{v}.\fld{blocks}$
    \State $\langle T_1, T_2\rangle \assign \Call{Split}{T,i}$
    \State \Call{CAS}{\var{v}.\fld{blocks}, $T$, $T_2$}
    \If{\var{v} is an internal \typ{Node}}
        \State \Call{Discard}{\var{v}.\fld{left}, \var{v}.\fld{blocks}[\var{i}].\eleft}
        \State \Call{Discard}{\var{v}.\fld{right}, \var{v}.\fld{blocks}[\var{i}].\eright}
    \EndIf
\EndFunction{Discard}
}
\end{algorithmic}
\caption{\label{pseudocode2GC}Queue implementation, continued. \here{Would this routine be simpler to write if i parameter was the rank within whole node rather than within the block?}}
\end{figure}

>>>>

\Function{void}{Propagate}{\typ{Node} \var{v}}
    \State \linecomment{propagate blocks from \var{v}'s children to root}
    \If{\bf{not} \Call{Refresh}{\var{v}}} \label{firstRefreshGC}  \hfill \com\ double refresh
        \State \Call{Refresh}{\var{v}} \label{secondRefreshGC}
    \EndIf
    \If{$\var{v} \neq \var{root}$} \hfill \com\ recurse up tree
        \State \Call{Propagate}{\var{v}.\fld{parent}}
    \EndIf
\EndFunction{Propagate}

@@

\gc{    
	\If{ \var{v} is root and $\var{h} \mbox{ mod } p^2=0$}
	\State \var{max} \assign 0
	\For {$k \assign 1..p$}
		\State $\var{max} \assign \max(\var{max}, \var{last}[k])$
	\EndFor 
	\State \Call{Help}{}
	\State \Call{Discard}{$\var{root}, \var{max}-2$}
	\EndIf
	\State
}

@@

\end{algorithmic}
\end{minipage}
\caption{Bounded-space queue implementation.\label{pseudocode1GC}}
\end{figure}

@@

\gc{
\Function{boolean}{Propagated}{\typ{Node} \var{v}, \typ{int} \var{b}}
    \State \linecomment{check if $\var{v}.\fld{blocks}[b]$ has been propagated to \var{root}}
    \State \linecomment{Precondition:  $\var{v}.\fld{blocks}[b]\neq \nl$}
    \If {$\var{v} = \var{root}$} \Return{true}
    \Else
    	\State \fld{dir} \assign\ (\var{v}.\fld{parent.left} = \var{v} ? \fld{left} : \fld{right}) 
	    \State \var{sup} \assign\ \var{v}.\fld{blocks}[\var{b}].\fld{super}
	    \If{$\var{v}.\fld{parent.blocks}[\var{sup}] \neq \nl$ and $\var{b} > \var{v}.\fld{parent.blocks}[\var{sup}].\edir$} \var{sup} \assign\ $\var{sup}+1$
	    \EndIf

	    \State \Return{$\var{v}.\fld{parent.blocks}[\var{sup}] \neq \nl$ and $\var{b} \leq \var{v}.\fld{parent.blocks}[\var{sup}].\edir$ and \Call{Propagated}{\var{v}.\fld{parent}, \var{sup}}}
    \EndIf
\EndFunction{Propagated}

\Function{void}{Help}{}
    \State \linecomment{help pending dequeues that have been propagated to root}
    \For{\bf{each} $\ell$ in $\var{leaf}[1..k]$}
        \State $\var{h} \assign \ell.\head$
        \If{$\ell.\fld{blocks}[\var{h}].\fld{element} = \nl$ and \Call{Propagated}{$\ell, h$}}
            \State $\langle\var{b}, \var{i}\rangle \assign\ \Call{IndexDequeue}{\ell, h, 1}$
            \State \var{output} \assign\ \Call{FindResponse}{$b, i$} 
            \State $\ell.\fld{blocks}[\var{h}].\fld{response} \assign\ \var{output}$
        \EndIf
    \EndFor
\EndFunction{Help}

\Function{void}{Discard}{\typ{Node} \var{v}, \typ{int} \var{t}}
    \State \linecomment{discard blocks with index less than or equal to threshold $t$ from RBT \var{v}.\fld{blocks} and all subblocks of those blocks}
    \State \var{t\sub{left}} = \var{v}.\fld{blocks}[\var{t}].\eleft
    \State \var{t\sub{right}} = \var{v}.\fld{blocks}[\var{t}].\eright
	\For{\var{j} \assign 1 to $p$}
        \State $T \assign \var{v}.\fld{blocks}$
        \State $T'  \assign \Call{Split}{T,t}$ \Comment{$T'$ is the root of new RBT containing $T$'s blocks with $\var{index} > t$}
        \State \bf{exit when} \Call{CAS}{\var{v}.\fld{blocks}, $T$, $T'$}\label{splitCASGC}
    \EndFor
    \If{\var{v} is an internal \typ{Node}} \Comment{discard subblocks in descendants of $v$}
        \State \Call{Discard}{\var{v}.\fld{left}, \var{t\sub{left}}}
        \State \Call{Discard}{\var{v}.\fld{right}, \var{t\sub{right}}}
    \EndIf
\EndFunction{Discard}
}
\end{algorithmic}
\caption{\label{pseudocode2GC}Bounded-space queue implementation, continued. }
\end{figure}
